---
title: "07-functions"
author: "Catarina Pien"
date: "9/15/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions tutorial

## Set up some test data

```{r load-data}
airtemps <- c(212, 30.3, 78, 32)

celsius1 <- (airtemps[1]-32)*5/9
celsius2 <- (airtemps[2]-32)*5/9
celsius3 <- (airtemps[3]-32)*5/9
celsius4 <- (airtemps[4]-32)*5/9

```

## Create function for converting fahrenheit to celsius
```{r create function}
F_to_C <- function(F) {
  celsius <- (F-32) * 5/9
  return(celsius)
}
```

## Test function
```{r test function}
celsius5 <- F_to_C(airtemps[1])
celsius5 == celsius1
celsius_array <- F_to_C(airtemps)
```

Create function for converting celsius to fahrenheit and test
```{r}
C_to_F <- function(C) {
  fahrenheit <- C * 9/5 + 32
  return(fahrenheit)
}

fahrenheit <- C_to_F(celsius_array)
fahrenheit1 <- C_to_F(celsius_array[1])
fahrenheit == airtemps
```

Create function to return multiple values in a data frame
```{r}
convertTemps <- function(F) {
  celsius <- (F-32) * 5/9
  kelvin <- celsius + 273.15
  return(list(fahrenheit = F, celsius = celsius, kelvin = kelvin))
}

temps_df <- data.frame(convertTemps(seq(-100, 100, 10)))
```

Create custom theme
```{r}
custom_theme <- function(title_size = 12, base_size = 9, legendside = "right") {
  ggplot2::theme(
    axis.ticks = ggplot2::element_blank(),
    text = ggplot2::element_text(family = 'Helvetica', color = 'gray30', size = base_size),
    axis.title.x = ggplot2::element_text(title_size),
    axis.title.y = ggplot2::element_text(title_size),
    legend.position = legendside
  )
}
```

Use function to create a graph
```{r}
library(ggplot2)

ggplot(temps_df, mapping = aes(x = fahrenheit, y = celsius, color = kelvin)) + geom_point() + 
  custom_theme(15, 13, "left")
```

aes_string or using formula to pass for customizing the variables that are being plotted
```{r}
scatterplot <- function(df, point_size = 2, font_size=9) {
  ggplot(df, mapping=aes(x=fahrenheit, y=celsius, color=kelvin)) +
    geom_point(size=point_size) +
    custom_theme(font_size)
}

scatterplot(temps_df, 3, 16)
```

## Make a package
```{r}

```
#Christina is making edits
# Hi Christina and pup!

