---
title: "BayesianModeling"
author: "Catarina Pien"
date: "10/27/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rjags)
library(mcmcplots)
library(broom.mixed)
library(tidyverse)
```

## Write model for snow fence
Save as text file
```{r}
mscript <- "model{
  for(i in 1:N){
      # Likelihood
      y[i] ~ dpois(theta[i]*x[i])
      
      # Prior for theta
      theta[i]~dgamma(alpha, beta)
  }
  
  # Root node priors for alpha and beta
  alpha ~ dgamma(2,1)
  beta~dexp(1)
  
  # Calculated population mean and standard deviation
  pmean <- alpha/beta
  pstd <- sqrt(alpha/beta^2)

}"
```

## Define data
```{r}
# y = number of invasive species
# x = area of snow fence

sf <- data.frame(y = c(138, 91, 132, 123, 173, 124, 109, 154, 138, 134),
                 x = c(72, 50, 55, 60, 78, 63, 54, 70, 80, 68))

# data must be in list for JAGS. Must be named same as in model.
datlist <- list(y = sf$y, 
                x = sf$x,
                N = nrow(sf))
```

## Consider initial models
* Need to give initials (starting values) to the root nodes, or functions with numbers in them
* Depends on number of chains
```{r}
# alpha = population rate of invasive plants
# Calculate sample mean to get an idea of where to start:
mean(sf$y/sf$x)

# 0.5, 5, 20 are target rates (theta). rate = alpha/beta
# 3 chains from different starting places. You want these chains to converge.
initslist <- list(list(alpha = 0.5, beta = 1),
                  list(alpha = 5, beta = 1),
                  list(alpha = 40, beta = 2))
```

## Compile model
```{r}
jm <- jags.model(file = textConnection(mscript),
                 data = datlist,
                 inits = initslist,
                 n.chains = 3)

# Do it another 1000x
update(jm, 1000)
```

## Get posterior (coda)
```{r}
jm_coda <- coda.samples(model = jm,
                        variable.names = c("theta", "alpha", "beta", "pmean", "pstd"), #Which variables do you want to monitor?
                        n.iter = 3000) # total will be 9000  models (3000*3) 
```

## Model diagnostics
Has it converged?
* Three colors overlap

Are there independent samples?
* Fuzzy is even, not going up and down together

Check autocorrelation plot
* Take every 20th sample there will be less correlation, but would have to run 20x more.


```{r}
traplot(jm_coda,
        parms=c("theta", "alpha", "beta"))

mcmcplot(jm_coda,
         parms = c("alpha", "beta"))
```

## Rerun with thinning by 20
```{r}
jm_coda2 <- coda.samples(model = jm,
                        variable.names = c("theta", "alpha", "beta", "pmean", "pstd"), #Which variables do you want to monitor?
                        n.iter = 3000*20,
                        thin = 20) # thin to only keep every 20 values
```

```{r}
mcmcplot(jm_coda2,
         parms = c("alpha", "beta"))
```

## Rhat, or Gelman diagnostic
* parameter by parameter rhat
* Upper C.I. </=1.2 = converged
* If not converged, might run for longer, use where previous model left off for starting values. Could also redefine priors (limit). Model could also not be specified correctly.
```{r}
gelman.diag(jm_coda2,
            multivariate = FALSE)
```

## Caterpillar plots
* Rates for fences 1:10
* Snow fences 8 and 9 have statistically different means
```{r}
caterplot(jm_coda2, 
          parms = "theta",
          reorder = FALSE)

caterplot(jm_coda2, parms = "pmean",
          reorder = FALSE)
caterplot(jm_coda2, parms = "pstd",
          reorder = FALSE)


```

## Results - summary of posterior (and plot)
```{r}
summary(jm_coda2[[1]])
head(jm_coda2[[1]])

broom.mixed::tidyMCMC(jm_coda2,
                      conf.int = TRUE, 
                      conf.level = 0.95)%>%
  dplyr::filter(grepl("theta", term)) %>%
  ggplot2::ggplot(aes(x = term, y = estimate)) +
  geom_pointrange(aes(ymin = conf.low,
                      ymax = conf.high))

```

## Effective number of parameters
# penalty = effective number of parameters
```{r}
dic.samples(jm, n.iter = 3000)
```











# Example Activity

* m should be a dummy variable of 0 and 1
```{r}
mscriptMoth <- "model{

  for(i in 1:N){
      # Likelihood
      y[i] ~ dbin(p[i], n[i])
      
      # Prior for theta
      logit(p[i]) <- beta1*m[i] + beta2*x[i]
  }
  
  # Root node priors for betas
  beta1 ~ dnorm(0,1E-6)
  beta2 ~ dnorm(0,1E-6)
  
  # Calculated population mean and standard deviation
  pmean <- n*p
  pstd <- sqrt(n*p*(1-p))

}"
```


```{r}
moths <- data.frame(site = rep(1:7, each = 2),
                    morph = rep(1:2, times = 7),
                    distance = rep(c(0, 7.2, 24.1, 30.2, 36.4, 41.5, 51.2), each = 2),
                    placed = rep(c(56, 80, 52, 60, 60, 84, 92), each = 2),
                    removed = c(13, 14, 28, 20, 18, 22, 9, 16, 16, 23, 20, 40, 24, 39))

datlist2 <- list(y = moths$removed,
                x = moths$distance,
                N = nrow(moths),
                n = moths$placed,
                m = moths$morph)
```

```{r}
jmMoth <- jags.model(file = textConnection(mscriptMoth),
                 data = datlist2,
                 n.chains = 3)
```

## Get posterior (coda)
```{r}
jmMoth_coda <- coda.samples(model = jmMoth,
                        variable.names = c("p", "beta1", "beta2", "pmean", "pstd"), #Which variables do you want to monitor?
                        n.iter = 3000) # total will be 9000  models (3000*3) 
```

## Model diagnostics
```{r}
traplot(jmMoth_coda,
        parms=c("p", "beta1", "beta2"))

mcmcplot(jmMoth_coda,
         parms = c("beta1", "beta2"))
```

## Rerun with thinning by 15
```{r}
jmMoth_coda2 <- coda.samples(model = jmMoth,
                        variable.names = c("p", "beta1", "beta2", "pmean", "pstd"), #Which variables do you want to monitor?
                        n.iter = 3000*15,
                        thin = 15) # thin to only keep every 15 values

mcmcplot(jmMoth_coda2,
         parms = c("beta1", "beta2"))
```


```{r}
sum_out <- broom.mixed::tidyMCMC(jmMoth_coda2,
                      conf.int = TRUE, 
                      conf.level = 0.95)
beta_sum <- sum_out %>%
  filter(grepl("beta", term)) %>%
  ggplot(aes(x = term, y = estimate))+
  geom_hline(yintercept = 0) + 
  geom_pointrange(aes(ymin = conf.low,
                      ymax = conf.high))

beta_sum

```

## Caterpillar plots

```{r}
caterplot(jmMoth_coda2, 
          parms = "p",
          reorder = FALSE)

caterplot(jmMoth_coda2, parms = "pmean",
          reorder = FALSE)
caterplot(jm_coda2, parms = "pstd",
          reorder = FALSE)


```

## Results - summary of posterior (and plot)
```{r}
summary(jm_coda2[[1]])
head(jm_coda2[[1]])

broom.mixed::tidyMCMC(jm_coda2,
                      conf.int = TRUE, 
                      conf.level = 0.95)%>%
  dplyr::filter(grepl("theta", term)) %>%
  ggplot2::ggplot(aes(x = term, y = estimate)) +
  geom_pointrange(aes(ymin = conf.low,
                      ymax = conf.high))

```









