---
title: "Time-series-forecasting"
author: "Catarina Pien"
date: "11/1/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Modeling

### Create a model
```{r}
library(broom)
set.seed(103)

days <- 0:9
mass <- vector(length = 10)
for(t in seq_along(days)){
  mass[t] <- 1 + 2 * days[t] + rnorm(1, 0, 1)
}

linear_data <- data.frame(day = days, mass = mass)
plot(mass~day, 
     data = linear_data, 
     ylab = 'mass (g)', 
     ylim = c(1, 20),
     xlim = c(0, 15))
```

### Run model
```{r}
mod_linear <- lm(mass ~ 1 + days, data = linear_data)
tidy(mod_linear)
```

Look at coefficients
Need to check assumptions still
```{r}
mod_linear
```


### Predictions
Create new data frame
Generate vector of predictions
Plot of data - predicted next 5 days of growth
```{r}
newdays <- 11:15
newdat <- data.frame(days = newdays)
preds <- predict(mod_linear, 
                 newdata = newdat)

plot(mass ~ day, 
     data = linear_data, 
     ylab = 'mass (g)', 
     ylim = c(1, 30),
     xlim = c(0, 15)) +
  points(newdays, preds, col = 'red') +
  abline(coef(mod_linear)) + 
  abline(confint(mod_linear)[,1], lty = 2) +
  abline(confint(mod_linear)[,2], lty = 2)
```

## Time Series Objects in R

* White noise model
* Autocorrelation with 1 month lag
* Autocorrelation and a trend
* Seasonal patterns
* Seasonal pattern and a trend

```{r}
set.seed(210)
months <- 1:240
noise <- rnorm(length(months))

lag   <- vector() 
for(t in 1:length(months)){
  if(t == 1){
    lag[t]  <- rnorm(1)
  } else {
    lag[t]  <- (lag[t-1] + noise[t]) / 2
  }
}

lag_trend <- lag + months / 48

seasonal <- 2*sin(2*pi*months/12) + noise

seasonal_trend <- seasonal + months / 48 

```

Now lets create the multivariate time series object:
```{r}
all <- ts(data = data.frame(noise, lag, lag_trend, seasonal, seasonal_trend), 
          frequency = 12)

plot(all)
tsdisplay(all)
```

Which of these have a lag?
```{r}
lag.plot(all, set.lags = c(1, 3, 6, 9))
```

Autocorrelation plots
```{r}
acf(all[,'noise'], xlab = 'Lag (years)')

```

```{r}
acf(all[,'lag'], xlab = 'Lag (years)')
```

```{r}
acf(all[,'seasonal'], xlab = 'Lag (years)')
```


### Seasonal Decomposition

* random = mean 0, sd = 1
* trend = moving average
```{r}
dec <- decompose(all[,'seasonal_trend'])

plot(dec)
```


```{r}
dec_df <- data.frame(trend = dec$trend, month = months)
dec_df <- dec_df[!is.na(dec_df),]

tidy(lm(trend ~ month, data = dec_df))
```

### Seasonal Trend with loess

* less even seasonal trend, more useful for describing what is observed
```{r}
seasonal_stl <- stl(all[,'seasonal_trend'], s.window = 6)
plot(seasonal_stl)
```

### Forecasting Toolbox
```{r}
tmp <- data.frame(month = months, trend = seasonal_stl$time.series[,'trend'])
#plot(tmp$month, tmp$trend)
fit <- lm(trend ~ month, data = tmp)
coef(fit)
```

```{r}
tidy(fit)
```

### 
```{r}
library(forecast)
ts_fit <- tslm(all[,'seasonal_trend'] ~ trend + season, data = all)
summary(ts_fit)

```

```{r}
plot(forecast(ts_fit, h = 20))
```

```{r}
ts_fit <- stlf(all[,'seasonal_trend'])
plot(forecast(ts_fit))

```

```{r}
summary(ts_fit)
```

```{r}
plot(forecast(ts_fit, h = 20))
```

### Playing with data
```{r}
library(tidyverse)
library(daymetr)
library(lubridate)
mac_daymet_list <- download_daymet(site = "Maricopa Agricultural Center",
                                   lat = 33.068941,
                                   lon =  -111.972244,
                                   start = 2000,
                                   end = 2020, internal = TRUE)

mac_daymet <- mac_daymet_list$data %>% 
    transmute(date = ymd(paste0(year, '01-01'))+ days(yday) -1, 
              precip = prcp..mm.day., 
              tmax = tmax..deg.c.,
              tmin = tmin..deg.c.,
              tmean = (tmax + tmin) / 2,
              trange = tmax - tmin,
              srad = srad..W.m.2.,
              vpd = vp..Pa.) 

readr::write_csv(mac_daymet, file = '../mac_daymet.csv')
```

Read in data
```{r}
mac_daymet <- readr::read_csv('../mac_daymet.csv') %>% 
    select(date, precip, tmean, srad, vpd)

head(mac_daymet)
```

```{r}
# mean temperature
tmean.ts <- ts(mac_daymet$tmean, 
               start = c(2000, 1), 
               end = c(2020, 365), 
               deltat = 1/365)
tmean.ts

head(mac_ts)
mac_ts <- ts(mac_daymet, 
               start = c(2000, 1), 
               end = c(2020, 365), 
               deltat = 1/365)
```

```{r}
plot(tmean.ts, ylab = "Daily mean T", xlab = "Year")
```

### Correlation temperature with 1 day ago, 10 days ago, 180 days (6 months), 360 days (1 year).
* 100 days = next season
```{r}
lag.plot(tmean.ts, set.lags = c(1, 10, 100, 180, 360))
plot(acf(tmean.ts, lag.max = 365 ))
```

Look at monthly lag
```{r}
tmean_mo <- mac_daymet %>% 
    mutate(year = year(date), month = month(date)) %>% 
    group_by(year, month) %>% 
    summarise(tmean = mean(tmean), .groups = 'keep') %>% 
    ungroup() %>% 
    select(tmean)

tmean.mo.ts <- ts(tmean_mo, start = c(2000, 1), end = c(2020, 12), frequency = 12)

lag.plot(tmean.mo.ts, lags = 12)
# lag 12 = 12 months
# 3 months and 9 months most non-correlated
```

Monthly means
```{r}
plot(acf(tmean.mo.ts))

```


Cross-correlation plot
# vpd = difference between how much vapor the air can hold and how much it does hold
```{r}
vpd_mo <- mac_daymet %>% 
    mutate(year = year(date), month = month(date)) %>% 
    group_by(year, month) %>% 
    summarise(vpd = mean(vpd), .groups = 'keep') %>% 
    ungroup() %>% 
    select(vpd)

lag.plot(vpd_mo, lags = 12)
acf(vpd_mo)

# cross-correlation, how well correlated across time
# How  well does vapor pressure deficit predict temperature?
ccf(tmean_mo, vpd_mo)

plot(mac_daymet$tmean, mac_daymet$vpd)
```



